---
params:
  update_date: FALSE
  run_chunk: FALSE
date: "`r source('_supp/helper.R'); newdate_func(params$update_date)`"
title: "`r totitle('construction of confidence interval for effect sizes with')` $2\\times 2$ `r totitle('table','')`"
author: "Anh"
output: 
  bookdown::html_document2:
    code_folding: hide
    number_sections: true
bibliography: ["_supp/citation.bib"]
link-citations: true
---

\newcommand{\bf}[1]{\boldsymbol{#1}}
\newcommand{\hat}[1]{\widehat{#1}}
\newcommand{\mm}[1]{\mathbb{#1}}
\newcommand{\bar}[1]{\overline{#1}}
\newcommand{\tp}[1]{{#1}^{\top}}

\def\E{\Bbb{E}}
\def\V{\Bbb{V}}
\def\P{\Bbb{P}}
\def\I{{\large\unicode{x1D7D9}}}
\def\indep{\perp\!\!\!\!\perp}
\newcommand{\overeq}[2]{\stackrel{#1}{#2}}
\def\epsilon{\varepsilon} 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
ttt = purrr::partial(totitle, icon = "")
```

::: {.watermark}
*DRAFT*
:::

---

Let us deem the following table 

```{r tab11}
tab_2t2 = tibble(
  ` ` = c("Group 1","Group 2", "Total"),
  Success = c("$n_{11}$","$n_{21}$","$n_{+1}$"),
  Failure = c("$n_{12}$","$n_{22}$","$n_{+2}$"),
  Total = c("$n_{1+}$","$n_{2+}$","N")
)
kableExtra::kbl(tab_2t2, caption = "$2\\times 2$ table")%>%
  kableExtra::kable_styling("striped")%>%
  kableExtra::footnote(
    general = "$\\mathbf{n} = \\{n_{11},n_{12},n_{21},n_{22}\\}$: the observed table. ",
    footnote_as_chunk = T
    )
```

The aim is to compare proportion of success between 2 groups. To do so, three effect sizes that are *risk difference, relative risk* and *odd ratio* will respectively be considered. 

# `r ttt('risk difference')`

Let $\pi_1$ and $\pi_2$ are proportion of success event of group 1 and group 2 and their estimators are
$$
\hat{\pi}_1 = \frac{n_{11}}{n_{1+}}, \text{ and} \quad \hat{\pi}_2 = \frac{n_{21}}{n_{2+}}.
(\#eq:eq11)
$$
Also, 
$$
\Delta = \pi_1-\pi_2
(\#eq:eq12)
$$
is difference between proprotion of group 1 and group 2 and its estimator is $\hat{\Delta} = \hat{\pi}_1-\hat{\pi}_2$.

Suppose we have data with 1 explanatory variable $\bf{x} \in \{0,1\}^n$ that indicates whether an individual patient is treated, $x_j=1$ if patient $j$ is treated and $x_j=0$ otherwise. response variable $\bf{y} \in \{0,1\}^n$ indicating efficacy of each patients, so $y_j=1$ if patient $j$ is recovered and $y_j=0$ otherwise.  

















<!-- ----------------------------------------------------------------------------- -->

::: {.shad}
# _References_ {-}
:::
