---
params:
  update_date: FALSE
date: "`r source('_supp/helper.R'); newdate_func(params$update_date)`"
title: "`r totitle('giới thiệu bayes')`"
output: 
  bookdown::html_document2:
    code_folding: hide
    scroll_highlight: yes
    number_sections: true
    toc_depth: 3
bibliography: ["_supp/citation.bib"]
link-citations: true
---

\newcommand{\v}[1]{\boldsymbol{#1}}
\newcommand{\hat}[1]{\widehat{#1}}
\newcommand{\mm}[1]{\mathbb{#1}}
\newcommand{\bar}[1]{\overline{#1}}
\newcommand{\tp}[1]{{\v{#1}}^{\top}}
\newcommand{\scr}[1]{\mathscr{#1}}

\def\E{\Bbb{E}}
\def\V{\Bbb{V}}
\def\P{\Bbb{P}}
\def\I{{\large\unicode{x1D7D9}}}
\def\indep{\perp\!\!\!\!\perp}
\def\epsilon{\varepsilon} 
\def\logit{\text{logit}}

::: {.right .shad}
[tham khảo tại @rei2019]
:::

::: {.watermark}
*DRAFT*
:::

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
options("scipen"=100)
source("_supp/helper.R")
library(tidyverse)
ttt = purrr::partial(totitle, icon ="")
# stackrel
```

# Gibbs sampling

Let us consider the following random samples
$$
\begin{align}
& Y_1, \dots, Y_n \stackrel{\text{iid}}{\sim} \mathcal{N}(\mu,\sigma^2), \quad \text{where} \\
& \mu \sim \mathcal{N}(\gamma,\tau^2), \text{ and }\quad \sigma^2 \sim \mathcal{InvGam}(a,b)
\end{align}
(\#eq:eq1)
$$
thus, we obtain conditional posterio distribution as follows
$$
\mu|\sigma^2,\v Y \sim \mathcal{N}\bigg(\frac{\tp Y\v 1_n(1/\sigma^2)+\gamma/\tau^2}{n/\sigma^2+1/\tau^2},\frac{1}{n/\sigma^2+1/\tau^2}\bigg)
(\#eq:eq2)
$$
$$
\sigma^2|\mu,\v Y \sim \mathcal{InvGam}\bigg(n/2+a, (1/2)(\v Y -\mu\v 1_n)^\top(\v Y -\mu\v 1_n)+b \bigg)
$$

```{r}
y <- c(2.68,1.18,-0.97,-0.98,-1.03)
iter = 25000
samples = matrix(rep(NA,iter*2), ncol = 2)%>% `colnames<-`(c("mean","sigma2"))

#initial values
mu = mean(y); sigma2 = var(y)
n = length(y)

# prior
gam = 0; tau = 100^2; a <- b<- .1

for(i in 1:iter){
  
P = n/sigma2 + 1/tau
M = sum(y)/sigma2 + gam/tau
mu = rnorm(1,M/P,1/sqrt(P))

A = n/2+a; B = sum((y-mu)^2)/2+b
sigma2 = 1/rgamma(1,A,B)
  
  samples[i,]<- c("mu" = mu, "sigma2" = sqrt(sigma2))
}
s = samples %>% as.data.frame()

plot(s$mean,s$sigma2)
```





# Metropolis-Hastings (MH) sampling

Ta xem xét phân bố sau $$
Y_i|\boldsymbol{\beta }\sim \mathcal{Poi}[N\exp(\beta_1+\beta_2i)]
$$ với $i = 1,\dots,4$, như vậy ta có hàm likelihood $$
f(\boldsymbol{Y}|\boldsymbol{\beta}) \propto \prod_{i=1}^4\exp\{-N\exp(\beta_1+\beta_2i)\}\exp\{Y_i(\beta_1+\beta_2i)\}
$$

<!-- ----------------------------------------------------------------------------- -->

::: {.shad}
# References {.unnumbered}
:::
